---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  library(geneConverter),
  library(data.table),
  library(kableExtra)
)
```

# geneConverter

<!-- badges: start -->
<!-- badges: end -->

The goal of geneConverter is to provides functions to convert gene synonyms to gene names and homologous gene 
names of several organism includes human.

## Installation

You can install the development version of geneConverter like so:

``` r
#install.packages("devtools)
library(devtools)
install_github("thekaplanlab/geneConverter")
```

## Example `orthoConverter`

Here is an example data frame including gene names in one column:

```{r, echo=FALSE}
example<-read.csv("example.csv")
example<-example[,-1]
```


```{r, echo=FALSE}
knitr::kable(example, "simple")
```



`orthoConverter` function converts gene synonyms (if any) in the selected column to gene names. This makes it easier to transform gene synonyms to gene names even when the column contains a mixture of gene synonyms and gene names. This function can also be used to be sure all elements are gene names and there is no gene synonym in the table.
Output is the same data table with all other columns preserved. If given data completely consisist of Ensembl ID's, please choose Ensembl version of that organism while using `ctype` argument. If you want there is a table which explains the abbreviation and creatures for `ctype` argument. You can examine it with `View(cre_table)`


## Usage `orthoConverter`

```{r, results='hide', message=FALSE}
df<-orthoConverter(example, "c.elegans_genes", ctype = celeg)
```

```{r, highlight=FALSE, echo=FALSE}
knitr::kable(df, "simple")
```


If `drop = TRUE` argument used, result will be like this: 

```{r, results='hide', message=FALSE}
df<-orthoConverter(example, "c.elegans_genes", ctype = celeg, drop = TRUE)
```

```{r, highlight=FALSE, echo=FALSE}
knitr::kable(df, "simple")
```


